Transparency of COVID-19-related research - Analysis

### 2022-03-16 covid_transparency_dataset_creation.Rmd
Changing all datasets and re-analyse
Due to a small error in the search, we did everything from the scratch again and created new accurate datasets. Then, I re-analysed the results. Results need to be more thorough BTW.

---
title: "Transparency indicators across the dentistry and oral health literature"
author: Eero Raittio, Ahmad Sofi-Mahmudi, Erfan Shamsoddin, Sergio E Uribe, Julie Farmer
date: "1/8/2022"
output: html_document
---

@@

```{r}
pacman::p_load(tidyverse,
               rtransparent, 
               metareadr, 
               europepmc)
```

@@

```{r}
COVID_19_keywords <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vS6ibaTIJ0-4MiOZrLHKwo-Ya3MjlGX4f-dPQwdmWwQZhRGi6C83K1I_WceUJPH9zGyBc4L-lSalET_/pub?output=csv")
```

@@

@@

db_2020_1 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2020-06-30]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")

db_2020_1 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2020-06-30]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")

db_2020_2 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-07-01 TO 2020-12-31]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")

db_2021_1 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2021-01-01 TO 2021-06-30]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")

db_2020_2 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2021-07-01 TO 2021-12-31]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")

db <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2021-12-31]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 1000000, output = "parsed")

db <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2021-12-31]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 1000000, output = "parsed")

db <- epmc_search(
  query = paste0(
    "'(",
    TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery,
    ") ",
    'AND (SRC:"MED") 
    AND (LANG:"eng" OR LANG:"en" OR LANG:"us") 
    AND (FIRST_PDATE:[2020-01-01 TO 2021-12-31])
    AND ((IN_EPMC:y) OR (OPEN_ACCESS:y))
    AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")',
    "'"
  ),
  limit = 100000, output = "parsed"
)
```

Let's see how many duplicates are there:
```{r}
table(duplicated(db$pmid))
```

Removing the duplicates:
```{r}
db <- db %>% distinct(pmid, .keep_all = TRUE)
```

Let's see the first six rows of the database:
```{r}
head(db)
```

Removing "PMC" from the cells
```{r}
db_2020_1$pmcid_ <- gsub("PMC", "", as.character(db_2020_1$pmcid))

db_2020_2$pmcid_ <- gsub("PMC", "", as.character(db_2020_2$pmcid))

db_2021_1$pmcid_ <- gsub("PMC", "", as.character(db_2021_1$pmcid))

db_2021_2$pmcid_ <- gsub("PMC", "", as.character(db_2021_2$pmcid))

db$pmcid_ <- gsub("PMC", "", as.character(db$pmcid))
```

Now, we make a folder for xml format articles and switch to that folder:
```{r}
dir.create("pmc")
setwd("pmc")
```

Next, we download xmls in format accessible with rtransparent:
```{r}
db_2020_1 <- subset(db_2020_1, pmcid_!=c(7264604, 7194244, 7323088, 7300787, 7276714, 7262104, 7228371))
db_2020_2 <- subset(db_2020_2, pmcid_!=c(7677857, 7164703, 7678408, 7486161, 7645956, 7486166, 7675701, 7405208, 7753328, 7575215, 7759253, 7568026, 7598225, 7416582, 7649313, 7492687, 7571788, 7675665, 7572235, 7457217))
db_2021_1 <- subset(db_2021_1, pmcid_!=c(8013335, 8013644, 8250125, 8014725, 8251275, 7486807, 8420411, 8251218, 8251305, 7894077))
db_2021_2 <- subset(db_2021_2, pmcid_!=8653067)

db <- subset(db_2021_2, pmcid_!=c(7264604, 8013335))

sapply(db_2020_1$pmcid_, mt_read_pmcoa)

sapply(db_2020_2$pmcid_, mt_read_pmcoa)

sapply(db_2021_1$pmcid_, mt_read_pmcoa)

sapply(db_2021_2$pmcid_, mt_read_pmcoa)

sapply(db$pmcid_, mt_read_pmcoa)

db <- subset(db, !(pmcid_ %in% c(8251275, 8251218, 7486161, 7228371, 8485571)))

@@

A list is created now. We should convert this list to a dataframe:
```{r}
df <- data.table::rbindlist(results_table_all, fill = TRUE)
```

@@

```{r}
setwd('..')
db <- db[!duplicated(db[, 4]),]
opendata <- merge(db, results_table_data, by = "pmid") %>% merge(df)
View(opendata)
```

>>>>

---
title: "Transparency indicators across the dentistry and oral health literature"
author: Ahmad Sofi-Mahmudi, Eero Raittio, Sergio E Uribe
date: "1/8/2022"
output: html_document
---

@@

```{r}
pacman::p_load(dplyr,
               rtransparent, 
               metareadr, 
               europepmc)
```

@@

```{r}
COVID_19_keywords <- read.csv("covid_keywords.csv")
```

@@

## Load all open-access and EPMC COVID-19-related papers indexed in PubMed

As in this step we should download a large file that includes the characteristics of papers of our query, to ease the process, we divided our query to five separate queries based on the date:
1- first half of 2020;
2- second half of 2020;
3- first half of 2021;
4- second half of 2021;
5- the first two months of 2022.

@@

db_2020_1 <- epmc_search(query = paste0("(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2020-06-30]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'), limit = 100000, output = "parsed")

total oa and non oa: 19467

db_2020_2 <- epmc_search(query = paste0("(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-07-01 TO 2020-12-31]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'), limit = 100000, output = "parsed")

total oa and non oa: 43107

db_2021_1 <- epmc_search(query = paste0("(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2021-01-01 TO 2021-06-30]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'), limit = 100000, output = "parsed")

total oa and non oa: 50512

db_2021_2 <- epmc_search(query = paste0("(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2021-07-01 TO 2021-12-31]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'), limit = 100000, output = "parsed")

total oa and non oa: 47975

db_2022 <- epmc_search(query = paste0("(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2022-01-01 TO 2022-02-28]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'), limit = 100000, output = "parsed")

total oa and non oa: 14272

## We then do the following for each of them

Let's see how many duplicates are there:
```{r}
table(duplicated(db_2020_1$pmid))
table(duplicated(db_2020_2$pmid))
table(duplicated(db_2021_1$pmid))
table(duplicated(db_2021_2$pmid))
table(duplicated(db_2022$pmid))

Removing the duplicates:
```{r}
db_2020_1 <- db_2020_1 %>% distinct(pmid, .keep_all = TRUE)
db_2020_2 <- db_2020_2 %>% distinct(pmid, .keep_all = TRUE)
db_2021_1 <- db_2021_1 %>% distinct(pmid, .keep_all = TRUE)
db_2021_2 <- db_2021_2 %>% distinct(pmid, .keep_all = TRUE)
db_2022 <- db_2022 %>% distinct(pmid, .keep_all = TRUE)

Creating a new column from pmcid column and removing "PMC" from the cells:
```{r}
db_2020_1$pmcid_ <- gsub("PMC", "", as.character(db_2020_1$pmcid))
db_2020_2$pmcid_ <- gsub("PMC", "", as.character(db_2020_2$pmcid))
db_2021_1$pmcid_ <- gsub("PMC", "", as.character(db_2021_1$pmcid))
db_2021_2$pmcid_ <- gsub("PMC", "", as.character(db_2021_2$pmcid))
db_2022$pmcid_ <- gsub("PMC", "", as.character(db_2022$pmcid))

Now, we make five folders for xml format articles:
```{r}
dir.create(c("pmc_2020_1", "pmc_2020_2", "pmc_2021_1", "pmc_2021_2", "pmc_2022"))
```

Next, we download xmls in format accessible with metareadr To skip errors (i.e., The metadata format 'pmc' is not supported by the item or by the repository.), first define a new function:

```{r}
skipping_errors <- function(x) tryCatch(mt_read_pmcoa(x), error = function(e) e)
```

Next, we download xmls in format accessible with rtransparent:
```{r}
setwd("pmc_2020_1")
sapply(db_2020_1$pmcid_, skipping_errors)

setwd("../pmc_2020_2")
sapply(db_2020_2$pmcid_, skipping_errors)

setwd("../pmc_2021_1")
sapply(db_2021_1$pmcid_, skipping_errors)

setwd("../pmc_2021_2")
sapply(db_2021_2$pmcid_, skipping_errors)

setwd("../pmc_2022")
sapply(db_2022$pmcid_, skipping_errors)

@@

A list is created now. We should convert this list to a dataframe:
```{r}
df1 <- data.table::rbindlist(results_table_all_f1, fill = TRUE)
df2 <- data.table::rbindlist(results_table_all_f2, fill = TRUE)
df3 <- data.table::rbindlist(results_table_all_f3, fill = TRUE)
df4 <- data.table::rbindlist(results_table_all_f4, fill = TRUE)
df5 <- data.table::rbindlist(results_table_all_f5, fill = TRUE)

@@

```{r}
setwd('..')
db <- db[!duplicated(db[, 4]),]
opendata <- merge(db, results_table_data, by = "pmid") %>% merge(df)
```

Random sample of search results for validation of our approach (100 refs):
```{r}
set.seed(100)

randomsample3 <- db3[sample(nrow(db3), 100), ]
```

### 2022-03-16 rmarkdown-covid-19-transparent-analysis.Rmd
Changing all datasets and re-analyse
Due to a small error in the search, we did everything from the scratch again and created new accurate datasets. Then, I re-analysed the results. Results need to be more thorough BTW.

# Results
Number of all papers (open access and non open access), open-access papers and open-access percentage:
```{r}
kable(data.frame(hits_all = 92752, 
           hits_oa = nrow(opendata),
           oa_percentage = round((nrow(opendata)/92752)*100, 2)))
```

@@

Characteristics of the paper with the highest number of citations:
```{r}
kable(opendata[which.max(opendata$citedByCount),] %>% 
  select(pmid, pmcid, doi.x, title, authorString, journalTitle, pubYear))
```

@@

First, importing jif file:
```{r}
jif <- read.csv("jif_2020.csv")
```

>>>>

# Results
Number of all papers (open access and non open access), open-access papers and open-access percentage:
```{r}
kable(data.frame(hits_all = 175550, 
           hits_oa = nrow(opendata),
           oa_percentage = round((nrow(opendata)/175550)*100, 2)))
```

@@

Characteristics of the paper with the highest number of citations:
```{r}
kable(opendata[which.max(opendata$citedByCount),] %>% 
  select(citedByCount, pmid, pmcid, doi.x, title, authorString, journalTitle, pubYear))
```

@@

First, importing jif file:
```{r}
jif <- read.csv(here("data", "jif_2020.csv"))
```
